---
title: "ST512-SP26-Homework-1"
author: Jason Menjivar-Hernandez
format: pdf
editor: visual
---

## Problem 1

n = 36

$\bar{x}$ = 23.4

sd (s) = 7.2

CI: $\bar{x} \pm$ t\* x SE

SE = $\frac{s}{\sqrt{n}}$

SE = $\frac{7.2}{\sqrt{36}}$ = 1.2

t\* = qt(1-(.04/2), 36) = 2.13

CI: 23.4 $\pm$ 2.13 \* 1.2

96% CI: (20.84,25.96)

## Problem 2

$H_o : \mu = 3.5$

$H_A: \mu < 3.5$

Will be conducting a left-tailed test because we want to know if the drug delivers relief faster (i.e., less than 3.5 minutes).

n = 50

$\bar{x}$ = 3.1

s = 1.5

$\alpha$ = 0.05

SE = $\frac{s}{\sqrt{n}} = \frac{1.5}{\sqrt{50}}$ = 0.212

t-stat = $\frac{\bar{x}-\mu_o}{SE}$ = $\frac{-0.4}{0.212}$ = -1.887

p-value = pt(t.value, df) = 0.033

p-value \< $\alpha$, we reject $H_o$

Therefore, there is sufficient evidence at the 5% confidence interval to conclude that the newly developed pain reliever delivers perceptible pain relief more quickly than the standard pain reliever.

## Problem 3

|              |          |          |
|--------------|----------|----------|
|              | Design 1 | Design 2 |
| $n_i$        | 11       | 6        |
| $\bar{y}_i.$ | 52       | 46       |
| sd           | 12       | 10       |

**a**

point estimate: $\hat{\mu_1 - \mu_2}$ = $\bar{y}_1.$ - $\bar{y}_2.$ = 52 - 46 = 6

95% CI: ($\bar{y}_1.$ - $\bar{y}_2.$) $\pm$ t\* x SE

Assuming equal variance:

df = $n_1 + n_2$ - 2 = 15

$s_p^2$ = $\frac{(n_1 - 1)s_1^2 + (n_2 - 1)s_2^2}{df}$ = $\frac{10(12^2)+5(10^2)}{15}$ = 129.33

$s_p = \sqrt{129.33}$ = 11.37

SE = $s_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}$ = 11.38$\sqrt{\frac{1}{11} + \frac{1}{6}}$ = 5.78

t\* = qt(0.975,15) = 2.13

95% CI: 6 $\pm$ 2.13 \* 5.78

95% CI: (-6.31, 18.31)

**b**

$H_o: \mu_1 = \mu_2$

$H_A: \mu_1 \neq \mu_2$

t-stat = $\frac{point estimate - 0}{SE}$ = $\frac{6}{5.78}$ = 1.04

p-value: 2 \* p(T \> \|t-value\|) = 0.31

$\alpha$ = 0.01

p-value \> $\alpha$; fail to reject $H_o$

There is not sufficient evidence at the 1% significance level to conclude the mean monthly sale differ between the two designs.

## Problem 4

**a**

This is an experimental design because the treatments are being controlled, the investigator's role is active

**b**

This is a randomized complete block design.

**c**

i.  response variable: tomato yield

ii. factor: fertilizer application rate

iii. treatments: 3, 5, and 7 units/acre

**d**

There are 6 replications per treatment.

## Problem 5

**a**

| **Source**         | **df** | **Sum of Squares** | **Mean Squares** |
|--------------------|--------|--------------------|------------------|
| **Between**        | 2      | 1533               | 766.6            |
| **Within (Error)** | 20     | 6241               | 312.0            |
| **Total**          | 22     | 7774               |                  |

**b**

Refer to R code at end for these values:

F-value = 2.457

p-value = 0.111

Our null hypothesis is that the mean phosphorous content does not differ among the three varieties. Using the calculated F-value, 2.457, the p-value was calculated to be 0.111. In comparison to a standard significance level of 5%, our p-value is larger. Therefore, failing to reject the null hypothesis. I think the three varieties have the same mean phosphorous content.

## Problem 6

| *Source of Variation* | *df* | *SS*   | *MS*  | *F-value* | *P-value* |
|-----------------------|------|--------|-------|-----------|-----------|
| **Between**           | 4    | 2.281  | 0.570 | 4.32      | 0.0039    |
| **With-in (error)**   | 60   | 7.949  | 0.132 |           |           |
| **Total**             | 64   | 10.230 |       |           |           |

Based on the ANOVA table, the calculated p-value is approximately 0.0039. The p-value is less than the significance level of 0.01, thus we reject the null hypothesis. Therefore, it appears to be a difference in means between the five diets. 



## R Code

```{r}
qt(1-(.04/2), 36) #problem 1 

pt(-1.887,49) #problem 2 

qt(0.975,15) #problem 3.a

2*(1-pt(1.04,15)) #problem 3.b
```

```{r}
# problem 5 r-code 
group <- factor(c(rep("A",8), rep("B",8), rep("C",7)))

value <- c(
  rep(72 + sqrt((8-1)*312/8), 4), rep(72 - sqrt((8-1)*312/8), 4),
  rep(68 + sqrt((8-1)*310/8), 4), rep(68 - sqrt((8-1)*310/8), 4),
  rep(55 + sqrt((7-1)*321/7), 3), rep(55 - sqrt((7-1)*321/7), 4)
)

data <- data.frame(group, value)
anova_model <- aov(value ~ group, data = data)
summary(anova_model)
```

```{r}
#problem 6 p-value 
1-pf(4.32,4,60)
```
